---
title: "Experiment runner SOC model calibration"
author:
  - Nicolas Pelletier (nicolas.pelletier@agr.gc.ca)
  - Francis Durnin-Vermette (francis.durnin-vermette@agr.gc.ca), 
  - Arumugam Thiagarajan (arumugam.thiagarajan@ec.gc.ca)
date: "2023-04-14"
output: html_document
---

## Load all packages
```{r message=FALSE, warning=FALSE}
library(tidyr)
library(caret)
library(dplyr)
library(sensitivity)
library(boot)
library(doParallel)
library(parallel)
library(foreach)
library(purrr)
library(tidyverse)
library(ggplot2)
library(lhs)
library(reshape2)
library(imputeTS)
```
## Set main working directory
```{r}
main.dir <- "C://Users//pelletierni//OneDrive - AGR-AGR//Documents//SOC modelling//soc-modelling-framework-NP//"
```

## Load data
```{r}
site_data <- read.csv(paste0(main.dir,"data//input data//site data//LTE_Master_beta_apr_18.csv"))
dsm_data<- read.csv(paste0(main.dir,"data//input data//site data//lte_soc_30_cm_dsm_expid.csv"))
climate.dir=paste0(main.dir,"data//input data//climate data//W9param_TablesCleaned")
parameter_bounds <- read.csv(paste0(main.dir,"data//input data//parameter data//ipcct2_parameters_gsa.csv"), stringsAsFactors = FALSE)
yield <- read.csv(paste0(main.dir,"data//input data//c_input//slyield_merged_final.csv"))
polyids_all <- read.csv(paste0(main.dir,"data//input data//c_input//exp_location_SLC_linkage.csv")) %>%
  select(Exp_ID,POLYID)
cinput <- read.csv(paste0(main.dir,"data//input data//c_input//cninput.csv"))
```

## Load functions
```{r}
source(paste0(main.dir,"scripts//Recent//Top30cm_SOC_estimator.R")) # Tool to get standardized SOC values for top 30 cm of soil
source(paste0(main.dir,"scripts//Recent//data preparation ipcct2.R")) # Data preperation function
source(paste0(main.dir,"scripts//Recent//ipcct2_SOCTier2Model.r")) # IPCC tier 2 steady state model
source(paste0(main.dir,"scripts//Recent//Spinnup IPCCT2.R")) # Model initiation for IPCCt2
source(paste0(main.dir,"scripts//Recent//sensitivity analysis IPCCT2.R")) # GSA function for IPCCt2
source(paste0(main.dir,"scripts//Recent//SIR IPCCT2.R")) # SIR function to obtain posterior distribution
```

## Create folder structure
```{r}
folders_needed=c(paste0(main.dir,"data//prepared data"),
                 paste0(main.dir,"data//prepared data//IPCCT2"),
                 paste0(main.dir,"results"),
                paste0(main.dir,"results//IPCCT2"),
                paste0(main.dir,"results//IPCCT2//spin up"))

for (i in folders_needed){
  ifelse(dir.exists(i),FALSE, dir.create(i))
}
```


## Convert all SOC measurements to top 30cm equivalent
```{r eval=FALSE, include=TRUE}
stocks.df <- get.stocks(site_data)
#table(stocks.df$Exp_ID)
correct.fac <- obtain.dsm.correction(stocks.df,dsm_data,merge.key="Exp_ID")
write.csv(correct.fac,paste0(main.dir,"top30_correction_factors_from_DSM_method.csv"))

results=top.SOC.estimate(stocks.df,correct.fac,30)
write.csv(results,paste0(main.dir,"data//prepared data//IPCCT2//modelled_stocks_top_30_cm.csv"),row.names = F)

## plot results
give.n <- function(x){
  return(c(y = median(x)*1.55, label = length(x))) 
   # experiment with the multiplier to find the perfect position
}
 
ggplot(data=results,aes(y=modelled_SOC,x=input_type))+
  geom_boxplot()+
  xlab("measurements available")+
  ylab("modelled SOC in top 30cm")+
  stat_summary(fun.data = give.n, geom = "text", fun = median,
               position = position_dodge(width = 0.75))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle=90))
```
# Data preperation for IPCCt2 model
## Site data preparation
```{r}
# Site data
site_data_ipcct2<- data.prep.ipcct2(site_data_raw=site_data,
                              cinput.df=cinput,
                              yield.df=yield,
                              poly.df=polyids_all,
                              save.interim.df=FALSE,
                              interim.df.file=NA)

write.csv(site_data_ipcct2,paste0(main.dir,"data//prepared data//IPCCT2//sites_prepared_ipcct2.csv"),row.names = F)
```
## Climate data preperation
```{r}
climate_data_ipcct2 <- data.prep.climate(climate.dir)

write.csv(climate_data_ipcct2,paste0(main.dir,"data//prepared data//IPCCT2//climate_prepared_ipcct2.csv"),row.names = F)
```

## Create the "initial" C stocks dataset (first C stock measurement at each StockID)
```{r}
stocks.df <-read.csv(paste0(main.dir,"data//prepared data//IPCCT2//modelled_stocks_top_30_cm.csv"))

initial.stocks <- stocks.df %>%
  group_by(site) %>%
  arrange(year) %>%
  slice(1)

write.csv(initial.stocks,paste0(main.dir,"data//prepared data//IPCCT2//initial_stocks_prepared_ipcct2.csv"),row.names = F)
```

Get initial C stocks for each pool (model initiation)
```{r}
# Read the dataset created previously
climate_data_ipcct2 <- read.csv(paste0(main.dir,"data//prepared data//IPCCT2//climate_prepared_ipcct2.csv"))
site_data_ipcct2 <- read.csv(paste0(main.dir,"data//prepared data//IPCCT2//sites_prepared_ipcct2.csv"))
initial.stocks<- read.csv(paste0(main.dir,"data//prepared data//IPCCT2//initial_stocks_prepared_ipcct2.csv"))

# Map of POLYID per site
poly.index=data.frame(table(POLYID=site_data_ipcct2$POLYID,site=site_data_ipcct2$site))
poly.index[poly.index$Freq>0,] -> poly.index
poly.index <- merge(poly.index,initial.stocks,by="site")

#Add columns where initial C stocks per pool will be saved
poly.index$init_active_frac=NA
poly.index$init_slow_frac=NA
poly.index$init_passive_frac=NA

write.csv(poly.index,paste0(main.dir,"data//prepared data//IPCCT2//poly_index.csv"),row.names = F)

# Default parameters for IPCCt2
params.default=data.frame(tillfac_FT = 3.036,
         tillfac_RT = 2.075,
         wfac_irri  = 0.775,
         k10        = 18.5,
         k20        = 4.9,
         k30        = 7.4,
         k40        = 0.209,
         k50        = 0.00689,
         f1         = 0.378,
         f2         = 0.368,
         f3         = 0.455,
         f5         = 0.0855,
         f6         = 0.0504,
         f7         = 0.42,
         f8         = 0.45,
         tmax       = 45,
         topt       = 33.69,
         plig       = 3)

ipcc2.init= initiate.ipcc2(poly.index,params=params.default)

write.csv(ipcc2.init,paste0(main.dir,"results//IPCCT2//spin up//initial_stocks_ipcct2.csv"),row.names = F)
```

# Start of the experiments

## experiment parameters
```{r}
general_name="test2"
date_on="April2023"

gsa_sample_size=1000
SIR_sample_size=100000
SIR_resample_size=1000

Experiment_name=paste0(general_name,"_",date_on,"_GSA ",gsa_sample_size,"_SIR ",SIR_sample_size,"To",SIR_resample_size)

folders_needed=c(paste0(main.dir,"data//prepared data//",Experiment_name),
                paste0(main.dir,"data//prepared data//",Experiment_name,"//IPCCT2"),
                paste0(main.dir,"results//",Experiment_name),
                paste0(main.dir,"results//",Experiment_name,"//IPCCT2"),
                paste0(main.dir,"results//",Experiment_name,"//IPCCT2//GSA"),
                paste0(main.dir,"results//",Experiment_name,"//IPCCT2//SIR"))

for (i in folders_needed){
 ifelse(dir.exists(i),FALSE, dir.create(i))
}
```

## 
```{r}
climate_data <- read.csv(paste0(main.dir,"data//prepared data//IPCCT2//climate_prepared_ipcct2.csv"))
site_data_IPCC2 <- read.csv(paste0(main.dir,"data//prepared data//IPCCT2//sites_prepared_ipcct2.csv"))
initial_c <- read.csv(paste0(main.dir,"results//IPCCT2//spin up//initial_stocks_ipcct2.csv"))
parameter_bounds <- read.csv(paste0(main.dir,"data//input data//parameter data//ipcct2_parameters_gsa.csv"), stringsAsFactors = FALSE)
stocks_data <- read.csv(paste0(main.dir,"data//prepared data//IPCCT2//modelled_stocks_top_30_cm.csv"))
poly.index <- read.csv(paste0(main.dir,"data//prepared data//IPCCT2//poly_index.csv"))

# Get mean annual temperature for each polygon (just to identify missing data later on)
climate.avg=climate_data %>% group_by(POLYID) %>% summarise(mean_temp=mean(tavg,na.rm=T))

# Calibration - validation split
site_data_driver <- site_data_IPCC2 %>%
  group_by(site,till,POLYID) %>%
  summarise (mean_c_input= mean(cinput,na.rm=T)) %>%
  merge(initial_c, by=c("site","POLYID"),all=T) %>%
  merge(climate.avg, by=("POLYID"))
  
## Subset sites where all info is available
site_data_driver <- site_data_driver %>% 
  filter(!is.na(mean_c_input) & !is.na(modelled_SOC) & !is.na(mean_temp) & !is.na(init_active_frac))

set.seed(123)
# Create for groups of Cinput - each represents a quartile of the distribution
site_data_driver$c_input_group <- ntile(site_data_driver$mean_c_input, 4)  

# Balanced split for different drivers
trainIndex <- caret::createDataPartition(c(site_data_driver$till,site_data_driver$c_input_group), p = .9, 
                                  list = FALSE, 
                                  times = 1)

calibration_sites <- site_data_driver[ trainIndex,]$site
evaluation_sites <- site_data_driver[-trainIndex,]$site

calibration_data <- site_data_IPCC2 %>%
  filter(site %in% calibration_sites)

evaluation_data <- site_data_IPCC2 %>%
  filter(site %in% evaluation_sites)

# Save to file
write.csv(calibration_data,file=paste0(main.dir,
                                       "data//prepared data//",
                                       Experiment_name,
                                       "//IPCCT2//site_data_calibration.csv"),row.names=F)
write.csv(evaluation_data,file=paste0(main.dir,
                                      "data//prepared data//",
                                      Experiment_name,
                                      "//IPCCT2//site_data_evaluation.csv"),row.names=F)
```

## Sensitivity analysis
```{r}
gsa_method="soboljansen"

return_si <- run.GSA.IPCCT2(main.dir=main.dir,
                            parameter_bounds=parameter_bounds,
                            calibration_data=calibration_data,
                            stocks_data=stocks_data,
                            climate_data = climate_data, 
                            initial_c=initial_c,
                            method=gsa_method,
                            sample_size=gsa_sample_size)

write.csv(return_si,paste0(main.dir,"results//",Experiment_name,"//IPCCT2//GSA//",gsa_method,"_results.csv"),row.names = F)

sensitiy.plot <- ggplot(data=return_si)+
  geom_bar(aes(x=SI,y=as.factor(params),fill=type),stat="identity",position="dodge")+
  theme_bw()+
  ylab("Parameter")+
  scale_fill_hue()

ggsave(plot = sensitiy.plot, paste0(main.dir,"results//",Experiment_name,"//IPCCT2//GSA//sensitivity_",gsa_method,".png"))

```

# Obtain a list of sensitive parameters
```{r}
return_si <- read.csv(paste0(main.dir,"results//",Experiment_name,"//IPCCT2//GSA//",gsa_method,"_results.csv"))

sensitive_params <- return_si %>% filter(uci>0.05 & type=="Total") %>% select(params)

new_params_bounds <- parameter_bounds %>% 
  mutate(cal=ifelse(Parameter %in% sensitive_params$params,1,0)) %>%
  mutate(lower=ifelse(cal==0,value,lower),
         upper=ifelse(cal==0,value,upper))
```

# SIR calibration
## Make the LHS sample
```{r}
#=================================================================================
# read prior distribution
# (Required columns: Parameter, value, lower, upper)
paramBounds <- parameter_bounds
#=================================================================================
# names of parameters that are allowed to vary
varSI       <- sensitive_params$params
nParams     <- length(varSI)
#=================================================================================
# LHS sampling for SIR
#=================================================================================
# sample size (1000000 samples were used in Gurung et al., 2020)
n <- SIR_sample_size
X1 <- randomLHS(n = n, k = nParams)
# transform standard uniform LHS to prior distribution 
Y1 <- matrix(0, nrow = nrow(X1), ncol = nParams)
for(i in 1:nParams){
  pos <- which(paramBounds$Parameter == varSI[i])
  lower <- paramBounds[pos, "lower"]
  upper <- paramBounds[pos, "upper"]
  Y1[, i] <- qunif(X1[, i], min = lower, max = upper)
}
X <- as.data.frame(Y1)
names(X) <- varSI
X <- cbind("SampleID" = 1:nrow(X), X)

names_X <- names(X)
for (i in 1:nrow(parameter_bounds)){
  param_on <- parameter_bounds[i,1]
  param_def <- parameter_bounds[i,2]
  if (param_on %in% names_X){} else 
    {X[[param_on]]<-param_def}
}

head(X)
```

## Perform SIR calibration on IPCCt2 model parameters
```{r}
calibration_data=read.csv(paste0(main.dir,
                                       "data//prepared data//",
                                       Experiment_name,
                                       "//IPCCT2//site_data_calibration.csv"))

posterior.distribution <- run.SIR.IPCCT2(main.dir=main.dir,
                         parameter.df=X,
                         stocks.df=stocks_data,
                         calibration.df=calibration_data,
                         climate_data=climate_data,
                         initial_c=initial_c,
                         sample_size=SIR_sample_size,
                         resample_size=SIR_resample_size)
```

## check posterior distributions
```{r}
###############

sir_table <- posterior.distribution %>%
  reshape2::melt(id.vars="SampleID") %>%
  group_by(variable)%>%
  summarise(`2.5%` = quantile(value,probs = 0.025),
                `25%` = quantile(value,probs = 0.25),
                median = quantile(value,probs = 0.50),
                `75%` = quantile(value,probs = 0.75),
                `97.5%` = quantile(value,probs = 0.975))


write.csv(sir_table ,paste0(main.dir,"results//",Experiment_name,"//IPCCT2//SIR//SIR_summary.csv"),row.names=F)
write.csv(posterior.distribution ,paste0(main.dir,"results//",Experiment_name,"//IPCCT2//SIR//SIR_posterior_distribution.csv"),row.names=F)

## Graphing:

prior <- X %>%
  reshape2::melt(id.vars="SampleID")
posterior <- posterior.distribution %>%
  reshape2::melt(id.vars="SampleID")

sir_graph <- ggplot() +
  geom_density(data = posterior, aes(value), col = NA, fill = "red", alpha = 0.2) +
  geom_density(data = prior, aes(value), col = NA, fill = "blue", alpha = 0.2) +
  facet_wrap(~variable, scales = "free", ncol = 3) +
  theme_bw() +
  theme(panel.border = element_rect(colour = "black", fill = NA), panel.grid.major = element_blank(),
        axis.title.x=element_blank(),panel.grid.minor = element_blank()) +
  scale_y_continuous(expand = c(0,0)) + scale_x_continuous(expand = c(0,0)) +
  theme(aspect.ratio=1)

print(sir_graph)
ggsave(plot = sir_graph, paste0(main.dir,"results//",Experiment_name,"//IPCCT2//SIR//SIR_distributions.png"))
```








